---
title: "CITE-seq: sample metadata and Harmony integration"
subtitle: "Amrute, Luo et al., Nature 635, 423-433 (2024) | doi:10.1038/s41586-024-08008-5"
output: html_document
---

**Purpose.** Adds sample-level clinical metadata, orders samples and HF etiologies, integrates samples with Harmony on the SCT PCA and estimates the number of informative PCs.

**Inputs.** `../normalized.rds` (from `02`), `./sample_metadata.csv`.

**Outputs.** `normalized_harmony.rds`

**Run order.** Upstream: `02_rna_protein_normalization.R`. Downstream: `04_wnn_clustering.R`.

Interactive analysis notebook: run chunk by chunk.

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(ggplot2)
library(harmony)
library(ArchR)
library(ggsci)
library(cowplot)
library(harmony)
```

```{r}
sample <- readRDS("../normalized.rds")
```

# Add sample Level meta-Data
```{r}
SampleInfo<-read.csv("./sample_metadata.csv",stringsAsFactors = FALSE)
sample$sample<-sub(".*-", "", sample$bc)  

meta<-sample@meta.data
meta1<-merge(meta, SampleInfo, by.x="sample",by.y="SampleID")
row.names(meta1)<-meta1$bc
meta2<-meta1[match(row.names(sample@meta.data),row.names(meta1)),]
sample@meta.data<-meta2
```

# Factor Sample Levels
```{r}
sample$sample <- factor(sample$sample, levels = c("sample13","sample28","sample32","sample33","sample34","sample41","sample42",
                                                  "sample4","sample5","sample8","sample39",
                                                  "sample1","sample2","sample12","sample17","sample27",
                                                  "sample6","sample7","sample9","sample15","sample29","sample30"), ordered = TRUE)
```

# Factor Condition
```{r}
sample$HF.etiology <- factor(sample$HF.etiology, levels = c("Donor", "ICM/AMI", "ICM", "NICM"), ordered = TRUE)
```

```{r}
VlnPlot(object = sample, features = "PC_1", group.by = "sample", pt.size = .0) + NoLegend()
VlnPlot(object = sample, features = "PC_1", group.by = "HF.etiology", pt.size = .0) + NoLegend()
```

```{r}
DefaultAssay(sample) <- "SCT"
```

```{r}
sample2 <- RunHarmony(sample, c("sample"), reduction = "sctpca", reduction.save = "harmony_rna", assay.use = "SCT")
```

```{r}
VlnPlot(object = sample2, features = "harmonyrna_1", group.by = "sample", pt.size = .0) + NoLegend()
VlnPlot(object = sample2, features = "harmonyrna_1", group.by = "HF.etiology", pt.size = .0) + NoLegend()
```

```{r}
pct <- sample2[["harmony_rna"]]@stdev / sum(sample2[["harmony_rna"]]@stdev) * 100
# Calculate cumulative percents for each PC
cumu <- cumsum(pct)
# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
n_pcs_to_use_RNA <- which(cumu > 90 & pct < 5)[1]
n_pcs_to_use_RNA
```


```{r}
saveRDS(sample2, "./normalized_harmony.rds")
```

















