---
title: "Visium: load, normalize and merge sections"
subtitle: "Amrute, Luo et al., Nature 635, 423-433 (2024) | doi:10.1038/s41586-024-08008-5"
output: html_document
---

**Purpose.** Loads the 28 Visium sections listed in `metadata/spatial_samples.csv`, adds section metadata (patient, zone, days after MI, disease), normalizes each with SCTransform and merges them.

**Inputs.** `./converted_for_seurat/*.h5seurat`; `metadata/spatial_samples.csv`.

**Outputs.** `spatial_merged_normalized.rds`

**Run order.** Upstream: none (spatial data from Kuppe et al., Nature 2022). Downstream: `02_spatial_integrate.R`.

Interactive analysis notebook: run chunk by chunk.

```{r setup}
## ---- Shared helpers, paths and output folders (EDIT) ----
repo_dir <- ".."   # EDIT: root of this repository
source(file.path(repo_dir, "R", "utils.R"))
```

```{r}
library(Seurat)
library(dplyr)
library(harmony)
library(SeuratDisk)
```

```{r}
# Load and SCTransform every section listed in metadata/spatial_samples.csv;
# section metadata (sample, patient, zone, days after MI, disease) comes from the sheet
sheet <- read_sample_sheet(file.path(repo_dir, "metadata", "spatial_samples.csv"))
sections <- load_visium_samples(sheet, data_dir = "./converted_for_seurat")
```

```{r}
sample <- merge(sections[[1]], y = sections[-1])

DefaultAssay(sample) <- "SCT"
# Union of the per-section variable features, in section order
VariableFeatures(sample) <- unlist(lapply(sections, VariableFeatures), use.names = FALSE)
```

```{r}
saveRDS(sample, "./spatial_merged_normalized.rds")
```

```{r}

```


