---
title: "Visium: prediction scores and fibroblast-state plots for one section"
subtitle: "Amrute, Luo et al., Nature 635, 423-433 (2024) | doi:10.1038/s41586-024-08008-5"
output: html_document
---

**Purpose.** Maps an acute MI section onto the snRNA-seq atlas and plots cell-type prediction scores, FAP expression and fibroblast-state z-scores in space.

**Inputs.** `./converted_for_seurat/AKK002_157782_RZ_P3.h5seurat`; annotated snRNA-seq atlas.

**Outputs.** Spatial feature plots.

**Run order.** Upstream: `04`. Downstream: none.

Interactive analysis notebook: run chunk by chunk.

```{r setup}
## ---- Shared helpers, paths and output folders (EDIT) ----
repo_dir <- ".."   # EDIT: root of this repository
source(file.path(repo_dir, "R", "utils.R"))
snhf_atlas_rds <- "path/to/snHF_atlas.rds"   # EDIT: annotated snRNA-seq HF atlas (reference)
```

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(SeuratDisk)
library(ArchR)
library(Nebulosa)
```

```{r}
Convert("./converted_for_seurat/AKK002_157782_RZ_P3.h5ad", dest = "h5seurat", overwrite = TRUE)
spatial <- LoadH5Seurat("./converted_for_seurat/AKK002_157782_RZ_P3.h5seurat")
spatial <- SCTransform(spatial, assay = "RNA", verbose = FALSE, method = "glmGamPoi", vars.to.regress = "percent.mt")
```

# Ref Map to Human Global
```{r}
sample <- readRDS(snhf_atlas_rds)
```

# Ref Map
```{r}
anchors <- FindTransferAnchors(
  reference = sample,
  query = spatial,
  normalization.method = "SCT",
  reference.reduction = "harmony",
  dims = 1:50
)

predictions <- TransferData(anchorset = anchors, refdata = sample$cell.type, dims = 1:50)
spatial <- AddMetaData(spatial, metadata = predictions)
```

# Prediction Scores
```{r}
features = c("prediction.score.Lymphatic","prediction.score.Pericyte","prediction.score.TNKCells","prediction.score.Fibroblast","prediction.score.Endothelium","prediction.score.Endocardium","prediction.score.Myeloid","prediction.score.SMC","prediction.score.Adipocyte","prediction.score.Cardiomyocyte","prediction.score.Glia","prediction.score.Epicardium","prediction.score.Mast","prediction.score.max")
```

```{r}
FeaturePlot(spatial, features = c("prediction.score.Pericyte","prediction.score.TNKCells","prediction.score.Fibroblast","prediction.score.Endothelium","prediction.score.Endocardium","prediction.score.Myeloid","prediction.score.SMC","prediction.score.Adipocyte","prediction.score.Cardiomyocyte","prediction.score.Epicardium","prediction.score.max"),keep.scale="all", reduction = "spatial", cols = c("lightgrey", "darkred"), ncol = 5) & theme(plot.title = element_text(size = 10)) + NoLegend()
```

```{r}
plot_density(spatial, features = "prediction.score.Epicardium", reduction = "spatial") + NoLegend()
plot_density(spatial, features = "prediction.score.Endocardium", reduction = "spatial") + NoLegend()
```


```{r}
DefaultAssay(spatial) <- "prediction.scores"
p_list <- plot_density(spatial, c("prediction.score.Myeloid", "prediction.score.Fibroblast"), joint = TRUE, combine = FALSE, reduction = "spatial")
p_list[[length(p_list)]]
```

# FAP Expression
```{r}
DefaultAssay(spatial) <- "SCT"
FeaturePlot(spatial, reduction = 'spatial', features = "FAP") + scale_color_gradientn(colors=paletteContinuous("solarExtra"), oob=scales::squish, limits=c(0,1)) + NoLegend()

FeaturePlot(spatial, reduction = 'spatial', features = "POSTN") + scale_color_gradientn(colors=paletteContinuous("solarExtra"), oob=scales::squish, limits=c(0,5)) + NoLegend()
```

# Ground State
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("PI16","CCDC80", "FBLN1", "SVEP1")
spatial@meta.data$groundStatefib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "groundStatefib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# Myofib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("ACTA2","TAGLN")
spatial@meta.data$myofib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "myofib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# CCL2/THBS1
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("CCL2","CCL11","THBS1")
spatial@meta.data$CCL2fib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "CCL2fib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# APOE Fib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("APOE","AGT","RGS5")
spatial@meta.data$APOEfib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "APOEfib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# DLK1 Fib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("DLK1","GPX3")
spatial@meta.data$DLK1fib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "DLK1fib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# PTGDS Fib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("PTGDS","GPC3")
spatial@meta.data$PTGDSfib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "PTGDSfib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# APOD Fib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("APOD","IGFBP5")
spatial@meta.data$APODfib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "APODfib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# GDF15 Fib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("ATF5","SLC3A2","TRIB3","GDF15")
spatial@meta.data$GDF15fib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "GDF15fib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# POSTN/FAP Fib
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("POSTN","COMP","FAP","COL1A1","THBS4","COL3A1")
spatial@meta.data$POSTNfib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "POSTNfib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# IFN
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("ISG15", "MX1")
spatial@meta.data$IFNfib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "IFNfib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# PLA2G2A
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("PLA2G2A", "CFD")
spatial@meta.data$PLA2G2A_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "PLA2G2A_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# PCOLCE2
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("PCOLCE2", "IGFBP6")
spatial@meta.data$PCOLCE2fib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "PCOLCE2fib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```

# PRG4
```{r}
DefaultAssay(spatial) <- "SCT"
expdata <- GetAssayData(spatial)
Pop1 <- c("PRG4", "CXCL14")
spatial@meta.data$PRG4fib_z <- gene_set_zscore(expdata, Pop1)
FeaturePlot(object=spatial, features = "PRG4fib_z",pt.size=.5, reduction = 'spatial') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,2)) + NoLegend()
```




